---
title: "markdown"
format: html
params:
  region: "NWFSC"
---

```{r}
#| eval: false
#| warning: false
#| echo: false
params$region <- "NWFSC"
```

## Markdown Notation

Efficiently using this package in the stock assessment workflow requires a basic understanding of markdown notation. Google docs easily recognizes these types of calls and formats it during writing. 

Any in-text markdown within google docs can be brought into the local template using the function `update_local()`. This reference is intended to help the user smoothly incorporate markdown language into plain text so the features of `ASAR` can easily incorporate into the workflow.

In order to add a piece of code into the text directly encase R code within a pair of backticks (`). For example:

```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| eval: false

params$region
```


The above example shows how one would incorporate the *region* parameter into a sentence in a markdown or quarto file. In google docs, this notation formatted to green and the font roboto mono. 

If we definited the parameter of region as NWFSC, the result would render as `r params$region`

Other ways to incorporate markdown notation include:

  - in-line calculations: \`r 2+2` 
  resulting in `r 2+2`

