---
title: "Increasing Report Accessibility"
output:
  html_document:
    toc: true
vignette: >
  %\VignetteIndexEntry{Increasing Report Accessibility}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

When someone reads your document, they may read it with a [screen reader](https://www.afb.org/blindness-and-low-vision/using-technology/assistive-technology-products/screen-readers). To ensure that they are able to interpret your document as intended, the document must contain some essential features, like tagged elements, alternative text, and metadata.

We have built `asar` so that the produced documents uphold several [federal accessibility standards](https://www.section508.gov/test/documents/). However, to achieve our goal of producing accessible documents, **you must complete a few tasks yourself**. But don't fret! We will help you. Below, we provide several resources to help you achieve this important goal.

# Your to-do list

1.  **Check the accuracy** of each figure's alternative text ("alt text" for short).

    -   `asar`'s figures already contains some prewritten alt text that includes data from the model results file. This means that most of the work has been done for you! However, it is crucial that you check this information for accuracy and update it where necessary. This is especially true if the default figures have been modified.

    -   If you see text that looks like a placeholder (e.g., "The x axis, showing the year, spans from B.start.year to B.end.year..."), that means that there was at least one instance where our tool failed to extract a specific value from the model results, calculate a key quantity (like the start year of a biomass plot- aka "B.start.year"), and substitute it into the placeholder. Learn how to manually add the right values in section "How to edit your report's alt text and captions", below.

    -   Again, we stress that while we have extracted key quantities as accurately as possible, **we cannot guarantee that each quantity will have been calculated perfectly. Input data varies widely. It’s always your responsibility to check the accuracy of your figures' alt text.**

2.  **Write the final component** of each figure's alt text.

    -   This prewritten alt text usually contains 3/4 essential ingredients for well-written alt text. The remaining ingredient (#4): *the relationship between the variables shown (i.e., what the figure is conveying).* Since we can't program `asar` to analyze the figure's meaning, *you must provide this*.

# Guidance and resources

## Alternative text

### What is alternative text?

After using `create_template()` to create a skeleton of your document, you will see chunks containing `fig-alt:` in the figures.qmd files, like this:

````{verbatim}
```{r}
#| eval: false
#| fig-alt: '...'
```
````

The [`fig-alt`](https://quarto.org/docs/computations/execution-options.html#caption-and-alt-text) parameter in this chunk signifies that this is where you should add a description of your figure that can be read aloud by the screen reader. This description, otherwise known as [alternative text](https://www.section508.gov/create/alternative-text/) should answer this essential question:

> What is this image conveying?

### Which information belongs in alt text?

While tempting, tools like AI cannot be used to easily answer this question. Additionally, one should not use the caption as the alt text. Here are four essential ingredients for well-written alt text, as described by Drs. Silvia Canelón and Liz Hare in their talk, "Revealing Room for Improvement in Accessibility within a Social Media Data Visualization Learning Community"[^1]:

[^1]: Canelón, Silvia, and Liz Hare. *Revealing Room for Improvement in Accessibility within a Social Media Data Visualization Learning Community*, csv,conf,v6, 7 May 2021, [spcanelon.github.io/csvConf2021/slides/](spcanelon.github.io/csvConf2021/slides/).

1.  Type of data visualization (e.g., scatterplot, line graph, box-and-whisker plot)
2.  Axis variables
3.  Range of the data
4.  The relationship between the variables shown (i.e., what the figure is conveying)

Dr. Hare stresses the importance of ingredient #4 by explaining, *"Don't waste my time with 1-3 if you aren't going to include 4. While some automatic alt text processes mine some of this information, I don't want to spend time building a mental model of the graph if I can't find out what the graph says.*[^2]*"*

[^2]: Hare, Liz. *Writing Meaningful Alt Texts for Data Visualizations in R*, R Ladies NYC, 10 Oct. 2022, [lizharedogs.github.io/RLadiesNYAltText/](lizharedogs.github.io/RLadiesNYAltText/).

Both presentations are great resources for learning about alt text and will help you as you craft your own alt text!

> Remember: The first three essential ingredients should already be present in your figures' prewritten alt texts! You just need to check them for accuracy and provide ingredient #4.

### Example of alt text

Here is an example of a figure with a caption and alt text. The caption is shown directly below the figure and is written in the chunk's options (`fig.cap=""`). The alt text is also included in the chunk's options (`fig.alt=""`) but is not shown unless the webpage is inspected with Developer Tools or it's extracted with a screen reader.

```{r, warning=FALSE, eval = TRUE, fig.align='center', fig.cap= "Tree circumference and age for 5 orange trees.", fig.alt="A line graph showing how tree circumference increases with age for a set of 5 orange trees. Age, shown on the x axis, is measured in days since 1968/12/31 and spans from 118-1582 days. Circumference, shown on the y axis, spans from 30-214 mm. All trees showed an increasing trend of trunk circumference with age, with each tree starting with a circumference of 30-33 mm at age 0 and ending with a circumference of 140-216 mm at age 1582. At age 1582, the tree with the largest circumference was tree 4, followed by trees 2, 5, 1, and 3."}
library(ggplot2)

orange <- as.data.frame(Orange)
orange <- orange |>
  dplyr::mutate(Tree = base::factor(Tree,
    levels = c(1, 2, 3, 4, 5)
  )) |>
  dplyr::rename(
    Age = age,
    Circumference = circumference
  )

ggplot2::ggplot(
  data = orange,
  aes(
    x = Age,
    y = Circumference,
    color = Tree
  )
) +
  ggplot2::geom_line(size = 1) +
  ggplot2::geom_point(size = 2) +
  ggplot2::scale_color_viridis_d() +
  ggplot2::xlim(0, NA) +
  ggplot2::ylim(0, NA) +
  ggplot2::theme_bw() +
  labs(
    x = "Age (days since 1968/12/31)",
    y = "Orange Tree Circumference (mm)"
  )
```

The figure's alt text is written as such:

*A line graph showing how tree circumference increases with age for a set of 5 orange trees. Age, shown on the x axis, is measured in days since 1968/12/31 and spans from 118-1582 days. Circumference, shown on the y axis, spans from 30-214 mm. All trees showed an increasing trend of trunk circumference with age, with each tree starting with a circumference of 30-33 mm at age 0 and ending with a circumference of 140-216 mm at age 1582. At age 1582, the tree with the largest circumference was tree 4, followed by trees 2, 5, 1, and 3.*

### How to write ingredient #4

> Ingredient #4 = the relationship between the variables shown (i.e., what the figure is conveying).

There is no one-size-fits-all approach for explaining what a figure is conveying. We've included some prompts, below, to get you started, but *you will need to go beyond these prompts to properly complete this task.*

#### Many figures

-   Describe the span of the 95% confidence interval at a meaningful x axis value.
-   Describe the meaning of the legend.

#### Line graphs

-   Describe where the line increases and decreases.
-   Describe where the line dips below the reference point (if present).

#### Kobe plots

-   Describe where most points fall (in the four quadrants).

### How to edit your report's alt text and captions

To edit your rda’s alt text, follow these steps:

1.  Open your report's 09_figures.qmd file.
2.  Run the first code chunk, which saves the filepath of your rda directory as an object (it has the label `"set-rda-dir-figs"`).
3.  Find the two code chunks associated with the figure you're interested (e.g., recruitment). Run the first chunk, which will have "setup" in the label (e.g., `"fig-recruitment-setup"`,`"fig-spawning_biomass-setup"`, etc.).
4.  **Add to the alt text by** pasting the existing alt text with a new string *within the chunk*. To do this, find your existing alt text object, which is an object named with the figure's topic and "alt_text" (e.g., `recruitment_alt_text`). Then, make an object (a string) containing your additional text (e.g., `new_alt_text`). Then, paste together the existing alt text object and your new text object. For example:

```{r, eval = FALSE}
# the original alt text for the recruitment figure
recruitment_alt_text

# the new text that will be added on to the recruitment figure's alt text
new_alt_text <- "This is my new alt text."

# add the new text to the old text
recruitment_alt_text <- paste0(recruitment_alt_text, new_alt_text)
```

5.  **Replace the alt text by** updating the original alt text object *within the chunk*. To do this, reassign the original alt text object as your new text object. For example:

```{r, eval = FALSE}
# the original alt text for the recruitment figure
recruitment_alt_text

# the new text that will replace the recruitment figure's alt text
new_alt_text <- "This is my new alt text."

# replace the old alt text with the new alt text
recruitment_alt_text <- new_alt_text
```

**NOTES**:

1.  Changes to your alt text will be saved within your 09_figures.qmd file, but not within the rda file itself. To directly edit the rda file's alt text or caption, assign a new value to the text you wish to change. For example, if your rda is called `rda` and you want to change the caption to "my new caption", you'd enter the following command: `rda[["cap"]] <- "my new caption"`. To change the alt text, you'd change "cap" to "alt_text" (e.g., `rda[["alt_text"]] <- "my new alt text"`.). Save the changes to the rda's file by entering the following command (in this example, our rda is called "biomass_figure.rda"): `save(rda, file = 'biomass_figure.rda')`.

2.  Edit figure and table captions with the same process. Just substitute mentions of alt text with caption or cap, depending on the context.

3.  As stated earlier, if you see text that looks like a placeholder (e.g., "The x axis, showing the year, spans from B.start.year to B.end.year..."), that means that there was at least one instance where our tool failed to extract a specific value from the model results and substitute it into the placeholder. Please make sure that your alt text and captions contain the expected values before moving forward with your report. Check out the inst/resources/captions_alt_text_template.csv file in the `stockplotr` package to view the template with placeholders. The same package's `write_captions()` function shows how values are extracted from the model results and substituted into the placeholders.

### More resources

Looking for more resources for writing alt text? Check out the [NOAA Library's website for creating accessible documents](https://library.noaa.gov/Section508/CreatingDocs).

## Acronyms and Glossary

When you use an acronym, it will be included in a glossary table that is automatically added to the end of your report. Below are directions for using acronyms and editing the glossary.

**Location**: The glossary ("report_glossary.tex") is located in your report folder.

**Order**: The glossary is sorted alphabetically. Case can differentiate entries (e.g., "M" (natural mortality) is different from "m" (meter(s)).)

**Structure**: Each acronym has its own line, structured like this:

`\newacronym{<"label" or "key">}{<"short form" or "acronym">}{<"long form">}`

- "label" (or "key"): The term written in the report body. The label links the short form (see below) to the glossary entry. In our glossary, labels are typically lowercase. *Examples: bcurrent, noaa*.

- "short form" (or "acronym"): The term actually shown when the report is rendered. The short form may have formatting applied to it. *Examples: \$B_{current}$* (which will render as $B_{current}$), *NOAA*.

- "long form": The meaning of the short form. *Examples: current biomass of stock, National Oceanic and Atmospheric Administration*.

Here are two entries using the examples mentioned above:

`\newacronym{bcurrent}{$B_{current}$}{current biomass of stock}`

`\newacronym{noaa}{NOAA}{National Oceanic and Atmospheric Administration}`

### Survey names

Many people refer to surveys with nicknames or "informal" names that do not match up with administrative documentation. For example, the "Atlantic Surf Clam & Ocean Quahog Dredge" is sometimes called the "clam survey". To facilitate communication in the reporting process, **we have added survey names to the glossary**.

*Informal survey names* (i.e., nicknames used to refer to surveys that are not correct) are listed in the "Acronym" column.

*Formal survey names* (i.e., the correct survey names) are listed in the "Meaning" column.

If you would like us to add more survey name pairs to our table, please make an [issue](https://github.com/nmfs-ost/asar/issues) or, even better, submit a [pull request](https://github.com/nmfs-ost/asar/pulls). Please see the [GitHub Docs' "Contributing to a project" page](https://docs.github.com/en/get-started/exploring-projects-on-github/contributing-to-a-project) for step-by-step guidance in making a pull request. Thank you!

### Editing the glossary

Remember that your changes will only be reflected in your report folder's glossary .tex file, not in the file stored in the `asar` package's inst/ folder. If you regenerate your report folder, that main file will overwrite your edited version unless you have indicated otherwise.

If you would like to contribute suggestions to the glossary, please open a pull request or [issue](https://github.com/nmfs-ost/asar/issues).

### How to remove a glossary entry

Delete the entire line in the file.

### How to add a glossary entry

First, ensure that your new acronym is not already duplicated in the glossary as this will cause an error upon rendering.

Find the logical location for your acronym based on alphabetical order. Make a new line and add the appropriate information for your acronym and its meaning, based on the structure discussed above.

### How to edit a glossary entry

You can edit any entry as needed.

### How to indicate that a word is an acronym

First, check if the glossary contains your acronym; if it doesn't, you can edit the file and add it yourself (see section above).

Once your acronym is in the glossary, go back to the location of the acronym in your report. Encase it in curly brackets ({}), with "\gls" preceding it. For example, to indicate "ABC" is an acronym, you would write `\gls{ABC}`. Use this notation each time you use the acronym in your text.

*Good news*: when using this notation, you never have to spell out the full meaning of the acronym upon its first usage, or even remember the location of its first usage! The CTAN `glossaries` package takes care of all of that.

Your text will look like this:

> This is the first instance of `\gls{ABC}`. Here, `\gls{ABC}` is used a second time.

And it will render like this:

> This is the first instance of Acceptable Biological Catch (ABC). Here, ABC is used a second time.

For specialized commands that enable capitalization, reference to acronyms' meanings, and more, check out the Command Summary (pg 46) within [the `glossaries` package's beginners' guide](https://ctan.math.illinois.edu/macros/latex/contrib/glossaries/glossariesbegin.pdf).

## Tables

Please do not place figures inside tables. Doing so makes it very difficult for software to properly identify the structure of (i.e., *tag*) the table and its inner elements.

# Status of accessibility features
*Date updated: 10/1/2025*

## NOAA institutional repository submission requirements

This section states the status of `asar` reports' accessibility features as compared to [the NOAA "Big 5" Section 508 compliance requirements](https://library.noaa.gov/Section508/QuickStart).

### 1. Tagged content 🚧

**Expectation:** The PDF is a tagged PDF.

**Status**: `add_tagging()` adds tags to *some* PDFs. Keep reading for an explanation.

#### What isn't working

##### Tables

**Report compilation will fail unless the report does not contain tables.** Generally, tagging tables is a known issue among developers. It's very challenging. In the meantime, we suggest two workarounds: 1) placing your tables in a separate document, or 2) adding your tables as raw LaTeX code inside the 08_tables.qmd file. We are working on adding option #2 as a feature in a future `asar` release.

#### What's functional

1. Figures
2. Figure captions
3. Lists
4. Other textual aspects of reports (e.g., headers, sections, parts)
5. Math

### 2. Bookmarks ✅

**Expectation:** Bookmarks are present on documents over 20 pages and illustrate the structure of the document.

**Expectation:** If a Table of Contents is present, bookmarks should reflect this.

**Status**: Both expectations are met.

### 3. Alternative Text ❌

**Expectation:** Alt text is present for all figures, charts, maps, etc.

**Status**: `add_alttext()` is currently failing to add alternative text to images, though *it was successful in the past*. Currently, it adds the figure's filepath as the alt text. We are investigating why this feature is failing.

### 4. Logical reading order ✅

**Expectation:** The reading order of the elements is logical and follows the flow of the document.

**Status**: This expectation is met.

### 5. Document properties ✅

**Expectation:** Title and Language are present.

**Status**: This expectation is met.

## Other accessibility considerations

This section states the status of `asar` reports' accessibility features that are not explicitly referenced by the NOAA IR submission "Big 5" list.

### HTML

HTML reports more readily pass several accessibility criteria.

### Adding tags and alt text in the same function

`add_accessibility()` is functional. It combines `add_tagging()` and `add_alttext()`, allowing the user to add tagging and alternative text at once.

### Glossary

The glossary is functional. 

## NOAA Central Library Accessibility Checklist: PDF Documents

This section states the status of `asar` reports' accessibility features as compared to the more extensive [NOAA Central Library's accessibility checklist for PDFs](https://library.noaa.gov/ld.php?content_id=61618926).

The checklist reflects the results from a report made with a basic, exemplary `asar` workflow (such as that in the README).

### General Accessibility

| Element | Pass | Fail | Not Applicable |
| :----------- | :-: | :-: | :---: |
| File name is concise and free of spaces or special characters. |X| | |
| Document properties for Title and Language are added. |X| | |
| Track changes have been accepted and/or turned off. All comments have been removed/resolved. |X| | |
| If there is a Table of Contents (TOC), it was created using the TOC Command. | | |X (The TOC was created externally from Adobe but are constructed such that TOC is inserted appropriately)|
| Page numbering codes are used (as opposed to manually entering page numbers). |X| | |
| Footnotes have been created via the Insert Footnote tool. | | |X (Footnotes are created externally from Adobe but are constructed such that footnes are inserted appropriately)|
| Links contain descriptive text informing the user of the context and content of the linked page. Avoid linking terms such as “Click Here” or “Website.” |X| | |
| All hyperlinks are working and linking to their intended destination. | X| | |
| All information that is conveyed through the use of color is also available without color. |X| | |
| All color for images, text, etc. avoids red-green contrast when possible. (For colorblind viewers). |X| | |

### Headings & Fonts

| Element | Pass | Fail | Not Applicable |
| :----------- | :-: | :-: | :---: |
| The document uses recommended font styles (Arial, Veranda, Times New Roman, Georgia, Calibri). |X| | |
| Text is easy to read against the background of the document (Recommended color contrast ratio of 4.5:1). |X| | |
| The document has been formatted using the style elements for Headings and is presented hierarchically (i.e., Heading 1 to Heading 2 leading into body text). |X| | |

### Image Requirements

| Element | Pass | Fail | Not Applicable |
| :----------- | :-: | :-: | :---: |
| The document should be free of background images, watermarks, and scanned images of text. |X| | |
| All images and graphics should appear crisp and clear. |X| | |
| All images and non-text elements have alternative text. ||X (We are working on fixing the erraneous alt text.) | |
| Multiple associated images on the same page are grouped together and alternative text for the images has been added. | | |X|
| Multilayered objects (such as charts or graphs) have been flattened into a single image and alternative text for the image has been added. | |X (We are working on fixing the erraneous alt text.)| |
| Complex images include descriptive text like captions. |X| | |

### Lists & Tables

As stated above, we are still working on solutions to properly tag tables.

| Element | Pass | Fail | Not Applicable |
| :----------- | :-: | :-: | :---: |
| All tables were made using the Insert Table option. | | |X (Tables will be created externally from Adobe but will be constructed such that tables are inserted appropriately)|
| No blank or merged cells are present in the table(s). | |X| |
| Data tables should have the first row designated as “Header Row” within the Table Properties. | |X| |
| Under Table Properties, “Allow row to break across pages” should not be checked. (This impacts reading order.) | |X| |
| Tables are labeled and/or described as necessary. | |X| |
| No layout tables are present in the document. | |X| |
| All lists have been appropriately tagged using all four tags: L, LI, Lbl, and LBody. |X| | |

### Bookmarks, Tagging & Table of Contents

| Element | Pass | Fail | Not Applicable |
| :----------- | :-: | :-: | :---: |
| All documents more than 20 pages have logical bookmarks. These should reflect the structure of the document and often time may mirror the Table of Contents (TOC). |X| | |
| Items in the TOC are linked appropriately, and page numbers are correct. |X| | |
| PDF tags have been added to the document (auto-tagging is acceptable, but may require touch-up work). |X| | |
| Reading order is logical and correct based on the PDF tags. |X| | |