### Add custom tables and figures {#sec-add-plots_coding}

In your 'report' folder, open your figures and tables docs (probably named something like '08_tables.qmd' and '09_figures.qmd'). Add the following code there.

**IMPORTANT**: The label in your captions/alt text csv must match the label in your plot's chunk options, **MINUS the "fig-" or "tab-"**. For example, the label "custom_table1" in the csv would match up properly with the "tbl-custom_table1" label in a code chunk.
The same rule applies for figures: "custom_figure1" in the csv would match up with "fig-custom_figure1" in the chunk options' label.

#### Tables

Follow the instructions in the appropriate section below based on your table classification: [reg_reg](#sec-rr), [reg_long](#sec-long), [wide_reg](#sec-wr), [wide_long](#sec-long), [ewide_reg](#sec-er), or [ewide_long](#sec-el).

For all table types, add a setup chunk to define your table and its caption:

````{verbatim}
```{r}
#| label: 'tab-setup-custom'
#| include: false

# Example dataframe
your_df <- data.frame(
  "x" = c(1, 2, 3),
  "y" = c(4, 5, 6)
)

# Example table
your_table <- gt::gt(your_df)

# Define your caption
your_cap <- "This is the caption for my custom table."
```
````

##### Regular tables (`reg_reg`) {#sec-rr}

*Portrait orientation, fits on one page.*

Add this chunk to display the table:

````{verbatim}
```{r}
#| label: 'tbl-custom1'
#| echo: false
#| warning: false
#| tbl-cap: !expr your_cap

your_table
```
````

##### Wide tables (`wide_reg`) {#sec-wr}

*Landscape orientation, fits on one page.*

Wrap the chunk in landscape braces:

````{verbatim}
::: {.landscape}

```{r}
#| label: 'tbl-custom-wide'
#| echo: false
#| warning: false
#| tbl-cap: !expr your_cap

your_table |>
  gt::tab_options(table.width = pct(100), table.layout = "auto") |>
  gt::cols_width(everything() ~ pct(20))
```

:::
````

##### Long tables (`reg_long` or `wide_long`) {#sec-long}

*Splits by row across multiple pages. `wide_long` uses landscape; `reg_long` uses portrait.*

You will need a separate chunk for each page of the split. For example, if your table requires 2 pages, repeat this chunk twice, changing the index in `grp_pull(1)` to `grp_pull(2)`.

Also, notice how `gt_split()` extracts subsets of rows from the table to ensure it fits on one page. This function is from the excellent `gt` package. There is a bug preventing the `gt` function from working as expected, so we filed an issue and submitted [a pull request that fixes the issue](https://github.com/rstudio/gt/pull/2132). Until it's resolved, we've added the fixed function to {asar} so users can access it via `asar::gt_split()`.

````{verbatim}
```{r}
#| label: 'tbl-custom-long-p1'
#| echo: false
#| tbl-cap: !expr paste0(your_cap, ' (Page 1 of 2)')

your_table |>
  gt::tab_options(table.width = pct(100), table.layout = "auto") |>
  gt::cols_width(everything() ~ pct(20)) |>
  asar::gt_split(row_every_n = 38) |> # Use 28 for wide_long
  gt::grp_pull(1) # Change this to 2 for the next page
```
````

> **NOTE**: For `wide_long`, wrap the chunk in `::: {.landscape}` braces. For example:

````{verbatim}
::: {.landscape}
# code chunk shown above
:::
````

##### Extra-wide tables (`ewide_reg`) {#sec-er}

*Landscape orientation, split by column across multiple pages.*

Because extra-wide tables are split by column, you must have already run [`export_split_tbls()`](https://github.com/nmfs-ost/asar/blob/main/R/export_split_tbls.R) to create the `_split.rda` file. Load that split file to access the list of tables:

````{verbatim}
```{r}
#| label: 'tbl-setup-ewide-split'
#| include: false
# Load the list of split tables generated by export_split_tbls()
load(file.path(tables_dir, "example_tab_table_split.rda"))
table_list_split <- table_list
```
````

Then, for each split table in the list, create a landscape chunk:

````{verbatim}
::: {.landscape}

```{r}
#| label: 'tbl-custom-ewide-c1'
#| echo: false
#| tbl-cap: !expr paste0(your_cap, ' (Column split 1 of 3)')

table_list_split[[1]] |>
  gt::tab_options(table.width = pct(100), table.layout = "auto") |>
  gt::cols_width(everything() ~ pct(20))
```

:::
````

##### Extra-wide, long tables (`ewide_long`) {#sec-el}

*Landscape orientation, split by both column AND row.*

This requires a nested approach using the `table_list_split` loaded above. You must create chunks for every combination of column-split and row-split.

````{verbatim}
::: {.landscape}

```{r}
#| label: 'tbl-custom-ewide-long-c1-r1'
#| echo: false
#| tbl-cap: !expr paste0(your_cap, ' (Column split 1 of 2, Row split 1 of 3)')

table_list_split[[1]] |>
  gt::tab_options(table.width = pct(100), table.layout = "auto") |>
  gt::cols_width(everything() ~ pct(20)) |>
  asar::gt_split(row_every_n = 28) |>
  gt::grp_pull(1)
```

:::
````

The next chunk would be identical except for a different label, caption, and would have `gt::grp_pull(2)` instead of `gt::grp_pull(1)`, for instance. When adding the second split table, add `table_list_split[[2]]` instead of `table_list_split[[1]]`, then change the label, caption, and `gt::grp_pull()` as needed.

##### Final step: page break

If you are adding multiple tables, insert a page break between them:

```{verbatim}
{{< pagebreak >}}
```

#### Figures

The workflow for figures is very similar to the workflow for [regular tables](#sec-rr): 1) add a chunk to define your figure, its caption, and its alternative text, and then 2) add a chunk to display the figure.

In your figures doc, add the following chunk to display the figure in your .qmd when it's rendered:

````{verbatim}
```{r} 
#| label: 'fig-custom1' # choose a label
#| echo: false
#| warning: false
#| fig-cap: !expr This is your figure caption.
#| fig-alt: !expr This is your figure alternative text.
# Import your data
your_df <- data.frame(
  "x" = c(1, 2, 3),
  "y" = c(4, 5, 6)
)

# Make your figure
your_figure <- ggplot2::ggplot(
  your_df,
  aes(
    x = x,
    y = y
  )
) +
  ggplot2::geom_point()

# Show your figure
your_figure
``` 
````

If you are adding another figure, add a page break first:

```{verbatim}
{{< pagebreak >}}
```
