### Add custom tables and figures {#sec-add-plots_rda}

In your 'report' folder, open your figures and tables docs (probably named something like '08_tables.qmd' and '09_figures.qmd'). Add the following code there.

#### Tables

In your tables doc, add your first R chunk (shown below). This chunk is used to establish your `tables_dir` folder and load the library needed to plot your tables (`gt`).

Click the following links to learn more about [cross-references](https://quarto.org/docs/computations/execution-options.html) (label) and [chunk options](https://quarto.org/docs/computations/execution-options.html) (echo, warning, etc.).

````{verbatim}
```{r} 
#| label: 'set-rda-dir-tbls-custom' 
#| echo: false
#| warning: false
#| include: false
library(gt)
tables_dir <- "C:/path/to/your/tables" # set the path to your tables_dir
``` 
````

Now, follow the instructions in the appropriate section depending on your table classification: **regular**, **wide**, **long**, **extra-wide**, or **extra-wide and long**.

##### Regular table (`reg_reg`)

Add the following chunk to import your table's rda and save objects containing the table and caption:

````{verbatim}
```{r} 
#| label: 'tbl-setup-custom1' # choose a label
#| echo: false
#| warning: false
#| include: false
# load your rda and save with a table-specific name, then delete "rda"
load(file.path(tables_dir, "example_tab.rda"))
example_table_rda <- rda
rm(rda)

# save table, caption as separate objects
example_table <- example_table_rda$table
example_cap <- example_table_rda$caption
``` 
````

Then, add this chunk to display the table:

````{verbatim}
```{r} 
#| label: 'tbl-custom1' 
#| echo: false
#| warning: false
#| tbl-cap: !expr example_cap

example_table
``` 
````

##### Wide table (`wide_reg`)

Import your rda using the same "Regular table" setup chunk above. Then, add [landscape braces](https://quarto.org/docs/authoring/article-layout.html#landscape-mode) and the display chunk:

```{verbatim}
::: {.landscape}
```

````{verbatim}
```{r} 
#| label: 'tbl-custom-wide1' # choose a label
#| echo: false
#| warning: false
#| tbl-cap: !expr example_cap

# the tab_options() and cols_width() lines format the table to fit the landscape page
example_table |>
  gt::tab_options(table.width = pct(100), table.layout = 'auto') |>
  gt::cols_width(everything() ~ pct(20))
``` 
````

```{verbatim}
:::
```

> **NOTE**: Landscape braces only work in [Quarto v1.6+](https://quarto.org/docs/blog/posts/2024-11-25-1.6-release/). Please read [this Discussion](https://github.com/nmfs-ost/asar/discussions/186) to learn how to check your Quarto version and update it if necessary.

##### Long table (`reg_long` or `wide_long`)

Import your rda using the "Regular table" setup chunk. Long tables are split by row using `gt_split()`. You will need one chunk per page (e.g., if you have 2 pages, repeat the display chunk twice).

> **NOTE**: For `wide_long` tables, wrap these chunks in `::: {.landscape}` braces. Landscape braces only work in [Quarto v1.6+](https://quarto.org/docs/blog/posts/2024-11-25-1.6-release/). Please read [this Discussion](https://github.com/nmfs-ost/asar/discussions/186) to learn how to check your Quarto version and update it if necessary.

````{verbatim}
```{r} 
#| label: 'tbl-custom-long1' # choose a label
#| echo: false
#| warning: false
#| tbl-cap: !expr paste0(example_cap, ' (1 of 2)')

example_table |>
  gt::tab_options(table.width = pct(100), table.layout = 'auto') |>
  gt::cols_width(everything() ~ pct(20)) |>
  asar::gt_split(row_every_n = 38) |> # Use 28 if wide_long
  gt::grp_pull(1)                    # Use 2 for the second chunk, etc.
``` 
````

##### Extra-wide table (`ewide_reg`)

For extra-wide tables, load the split rda. This file contains a list of tables and their associated caption specifiers.

````{verbatim}
```{r} 
#| label: 'tbl-setup-ewide1' # choose a label
#| echo: false
#| warning: false
#| include: false
# Load the split rda
load(file.path(tables_dir, "example_tab_table_split.rda"))

# Save the list and the caption specifiers
example_table_split <- table_list
example_cap_split <- names(example_table_split)
rm(table_list)
```
````

Wrap each split table in landscape braces. Repeat the chunk below for each table in your `table_list`:

```{verbatim}
::: {.landscape}
```

Then, add a chunk for each split table. Note how the caption is constructed using the specifiers:

````{verbatim}
```{r} 
#| label: 'tbl-custom-ewide1-part1' 
#| echo: false
#| warning: false
#| tbl-cap: !expr paste0(example_cap_split, ' (1 of 2)')

example_table_split[[1]] |>
  gt::tab_options(table.width = pct(100), table.layout = 'auto') |>
  gt::cols_width(everything() ~ pct(20))
```
````

```{verbatim}
:::
```

Duplicate the "tbl-custom-ewide1-part1" chunk above, changing the label, caption, and item pulled from the example_table_split object (e.g., `example_table_split[[2]]`) to add the rest of the split tables.

##### Extra-wide, long table (`ewide_long`)

For tables that are both extra-wide and long, you will use the same split rda import as above, but you will also apply the row-splitting logic to each chunk:

```{verbatim}
::: {.landscape}
```

````{verbatim}
```{r} 
#| label: 'tbl-ewide-long-c1-r1' 
#| echo: false
#| tbl-cap: !expr paste0(example_cap_split, ' (1 of 2 tables split by column, 1 of 3 tables split by row)')

example_table_split[[1]] |>
  gt::tab_options(table.width = pct(100), table.layout = 'auto') |>
  gt::cols_width(everything() ~ pct(20)) |>
  asar::gt_split(row_every_n = 28) |>
  gt::grp_pull(1)
``` 
````

```{verbatim}
:::
```

Duplicate the "tbl-ewide-long-c1-r1" chunk above, changing the label, caption, item pulled from the example_table_split object (e.g., `example_table_split[[2]]`), *and* the portion of the table split by row (e.g., `grp_pull(2)`) to add the rest of the split tables.

##### Final step: page break

If you are adding multiple tables, regardless of size, add a page break between them to ensure they render on separate pages:

```{verbatim}
{{< pagebreak >}}
```

#### Figures

The workflow for figures is very similar to the workflow for regular tables.

In your figures doc, add your first R chunk (shown below). This chunk is used to establish your figures_dir folder.

Click the following links to learn more about [cross-references](https://quarto.org/docs/computations/execution-options.html) (label) and [chunk options](https://quarto.org/docs/computations/execution-options.html) (echo, warning, etc.).

````{verbatim}
```{r} 
#| label: 'set-rda-dir-figs-custom' # you can set this label as you wish, as long as it's not repeated in any other R chunk in, for instance, the 08_tables.qmd or 09_figures.qmd files. Labels are used for cross-referencing in your report text.
#| echo: false
#| warning: false
figures_dir <- "C:/path/to/your/figures" # set the path to your figures_dir
``` 
````

Add the following chunk to your script to import your figure's rda and save objects containing the figure, caption, and alternative text:

````{verbatim}
```{r} 
#| label: 'fig-setup-custom1' # choose a label
#| echo: false
#| warning: false
#| include: false
# load your rda and save with a figure-specific name, then delete "rda"
load(file.path(figures_dir, "example_fig.rda"))
example_figure_rda <- rda
rm(rda)

# save figure, caption, and alternative text as separate objects
example_plot <- example_figure_rda$figure
example_cap <- example_figure_rda$caption
example_alt_text <- example_figure_rda$alt_text
``` 
````

Then, add this chunk to display the figure in your .qmd when it's rendered:

````{verbatim}
```{r} 
#| label: 'fig-custom1' # choose a label
#| echo: false
#| warning: false
#| fig-cap: !expr example_cap
#| fig-alt: !expr example_alt_text
example_plot
``` 
````

If you are adding another figure, add a page break first:

```{verbatim}
{{< pagebreak >}}
```
