### Add custom tables and figures {#sec-add-plots_coding}

In your 'report' folder, open your figures and tables docs (probably named something like '08_tables.qmd' and '09_figures.qmd'). Add the following code there.

#### Tables

Follow the instructions in the appropriate section below based on your table classification: reg_reg, reg_long, wide_reg, wide_long, ewide_reg, or ewide_long.

For all table types, add a setup chunk to define your table and its caption:

````{verbatim}
```{r}
#| label: 'tab-setup-custom'
#| include: false

# Example dataframe
your_df <- data.frame(
  "x" = c(1, 2, 3),
  "y" = c(4, 5, 6)
)

# Example table
your_table <- gt::gt(your_df)

# Define your caption
your_cap <- "This is the caption for my custom table."
```
````

##### Regular tables (`reg_reg`)

*Portrait orientation, fits on one page.*

Add this chunk to display the table:

````{verbatim}
```{r}
#| label: 'tbl-custom1'
#| echo: false
#| warning: false
#| tbl-cap: !expr your_cap

your_table
```
````

##### Wide tables (`wide_reg`)

*Landscape orientation, fits on one page.*

Wrap the chunk in landscape braces:

````{verbatim}
::: {.landscape}

```{r}
#| label: 'tbl-custom-wide'
#| echo: false
#| warning: false
#| tbl-cap: !expr your_cap

your_table |>
  gt::tab_options(table.width = pct(100), table.layout = 'auto') |>
  gt::cols_width(everything() ~ pct(20))
```

:::
````

##### Long tables (`reg_long` or `wide_long`)

*Splits by row across multiple pages. `wide_long` uses landscape; `reg_long` uses portrait.*

You will need a separate chunk for each page of the split. For example, if your table requires 2 pages, repeat this chunk twice, changing the index in `grp_pull(1)` to `grp_pull(2)`.

> **NOTE**: For `wide_long`, wrap the chunk in `::: {.landscape}` braces.

````{verbatim}
```{r}
#| label: 'tbl-custom-long-p1'
#| echo: false
#| tbl-cap: !expr paste0(your_cap, ' (Page 1 of 2)')

your_table |>
  gt::tab_options(table.width = pct(100), table.layout = 'auto') |>
  gt::cols_width(everything() ~ pct(20)) |> 
  asar::gt_split(row_every_n = 38) |> # Use 28 for wide_long
  gt::grp_pull(1)                    # Change this to 2 for the next page
```
````

##### Extra-wide tables (`ewide_reg`)

*Landscape orientation, split by column across multiple pages.*

Because extra-wide tables are split by column, you must have already run `export_split_tbls()` to create the `_split.rda` file. Load that split file to access the list of tables:

````{verbatim}
```{r}
#| label: 'tbl-setup-ewide-split'
#| include: false
# Load the list of split tables generated by export_split_tbls()
load(file.path(tables_dir, 'example_tab_table_split.rda'))
table_list_split <- table_list
```
````

Then, for each split table in the list, create a landscape chunk:

````{verbatim}
::: {.landscape}

```{r}
#| label: 'tbl-custom-ewide-c1'
#| echo: false
#| tbl-cap: !expr paste0(your_cap, ' (Column split 1 of 3)')

table_list_split[[1]] |>
  gt::tab_options(table.width = pct(100), table.layout = 'auto') |>
  gt::cols_width(everything() ~ pct(20))
```

:::
````

##### Extra-wide, long tables (`ewide_long`)

*Landscape orientation, split by both column AND row.*

This requires a nested approach using the `table_list_split` loaded above. You must create chunks for every combination of column-split and row-split.

````{verbatim}
::: {.landscape}

```{r}
#| label: 'tbl-custom-ewide-long-c1-r1'
#| echo: false
#| tbl-cap: !expr paste0(your_cap, ' (Column split 1 of 2, Row split 1 of 3)')

table_list_split[[1]] |>
  gt::tab_options(table.width = pct(100), table.layout = 'auto') |>
  gt::cols_width(everything() ~ pct(20)) |>
  asar::gt_split(row_every_n = 28) |>
  gt::grp_pull(1)
```

:::
````

##### Final step: page break

If you are adding multiple tables, insert a page break between them:

```{verbatim}
{{< pagebreak >}}
```

#### Figures

The workflow for figures is very similar to the workflow for regular tables.

In your figures doc, add the following chunk to display the figure in your .qmd when it's rendered:

````{verbatim}
```{r} 
#| label: 'fig-custom1' # choose a label
#| echo: false
#| warning: false
#| fig-cap: !expr This is your figure caption.
#| fig-alt: !expr This is your figure alternative text.
# Import your data
your_df <- data.frame(
  "x" = c(1, 2, 3),
  "y" = c(4, 5, 6)
)

# Make your figure
your_figure <- ggplot2::ggplot(
  your_df,
  aes(
    x = x,
    y = y
  )
) +
  ggplot2::geom_point()

# Show your figure
your_figure
``` 
````

If you are adding another figure, add a page break first:

```{verbatim}
{{< pagebreak >}}
```

You're almost done! Proceed to the last section ([Update report skeleton](#sec-update-skeleton)) to finish adding your custom figures to your report.
