### Add custom tables and figures {#sec-add-plots_rda}

This step will alter your figures and tables docs (probably named something like '08_tables.qmd' and '09_figures.qmd'), located in your 'report' folder.

As a reminder, this workflow assumes that you have tables and/or figures stored in rda files. If you don't have those rda files, please see the [Export your custom tables and figures](#sec-export-custom-figs-tables) section to create them.

**IMPORTANT**: The label in your captions/alt text csv must match the label in your plot's chunk options, **MINUS the "fig-" or "tab-"**. For example, the label "custom_table3" in the csv would match up properly with the "tbl-custom_table3" label in a code chunk. The same rule applies for figures: "custom_figure1" in the csv would match up with "fig-custom_figure1" in the chunk options' label.

#### Figures

To add figures, first ensure that your folder containing the figures is called "figures".

Then, there are only two arguments to fill out:

1.  `subdir`: The location where the *new* figures doc should be saved (we recommend your 'report' folder, so it overwrites the old, empty version)
2.  `figures_dir`: The location of your "figures" folder

```{r}
#| label: add-figures
#| eval: false
#| warning: false
#| messages: false
create_figures_doc(
  subdir = fs::path(getwd(), "report"), # indicates the new figures doc will be saved in your "report" folder, located in the working directory
  figures_dir = getwd() # indicates your "figures" folder is located in the working directory
                  )
```

In your figures doc, you'll notice the following structure:

**Chunk 1** will always create an object saving the location of your `figures_dir`.

Then, there will be at least two chunks for each figure. For each figure,

**Chunk 2** will:

-   load an rda containing a figure
-   give the rda a specific name
-   save the figure, caption, and alternative text as separate objects

**Chunk 3** will:

-   display the figure

#### Tables

Adding tables entails nearly the same process as described for figures. The only differences are:

1.  Tables will involve captions only, whereas figure require captions and alternative text
2.  Tables can be rotated or split across pages. See the [direct coding-in-qmd workflow (Option 3)](#sec-add-plots_coding) to learn about the logic within `create_tables_doc()` and why some tables are split into several code chunks.

To add tables to the {asar} tables doc:

```{r}
#| label: add-tables
#| eval: false
#| warning: false
#| messages: false
create_tables_doc(
  subdir = fs::path(getwd(), "report"),
  tables_dir = getwd()
  )
```