### Export your custom tables and figures {#sec-export-custom-figs-tables}

Next, export your tables and figures to rda files. This code was adapted from functions like `stockplotr::plot_biomass`.

#### Import pre-written captions and alternative text

Import your csv containing your captions and alternative text and save it as an object. Then, save the captions and alternative text as separate objects. For example:

```{r eval=F}
# Import captions and alternative text
my_text <- read.csv(file = "captions_alt_text.csv")

# Extract table caption
my_table_cap <- my_text |>
  dplyr::filter(
    label == "example_table",
    type == "table"
  ) |>
  dplyr::select(caption) |>
  as.character()

# Extract figure caption
my_figure_cap <- my_text |>
  dplyr::filter(
    label == "example_fig",
    type == "figure"
  ) |>
  dplyr::select(caption) |>
  as.character()

# Extract figure alternative text
my_figure_alt_text <- my_text |>
  dplyr::filter(
    label == "example_fig",
    type == "figure"
  ) |>
  dplyr::select(alt_text) |>
  as.character()
```

> **NOTE**: Your label must match the label associated with your table or figure in your csv containing the captions and alternative text. If it does not match, the label will not be extracted properly and added to your rdas.

#### Make rdas

Your rdas will contain your table or figure, caption, and alternative text (if it's a figure).

##### Tables

```{r eval=F}
rda <- list(
  "table" = your_table, # your table object
  "caption" = my_table_cap # your table caption
)
```

##### Figures

```{r eval=F}
rda <- list(
  "figure" = your_figure, # your figure object
  "caption" = my_figure_cap, # your figure caption
  "alt_text" = my_figure_alt_text # your figure alternative text
)
```

#### Make 'figures' and 'tables' folders (if needed)

Do you already have 'figures' and 'tables' folders? *Hint: both should be in your working directory.* If yes: Great! You can skip this step. If not, simply enter the code below to make a the folders.

```{r eval=FALSE}
dir.create(fs::path(getwd(), "figures"))
dir.create(fs::path(getwd(), "tables"))
```

#### Export your rdas

##### Tables

This will produce an rda called "example_tab.rda" and save it in your 'tables' folder.

```{r eval=FALSE}
save(rda,
  file = fs::path(
    getwd(), # this is your tables directory
    "tables",
    paste0(
      "example_tab.rda" # your table's label
    )
  )
)
```

##### Figures

This will produce an rda called "example_fig.rda" and save it in your 'figures' folder.

```{r eval=FALSE}
save(rda,
  file = fs::path(
    getwd(), # this is your figures directory
    "figures",
    paste0(
      "example_fig.rda" # your figure's label
    )
  )
)
```

