### Identify table width, length, orientation, and if splitting is necessary {#sec-id_width}

*Skip this section if you are not including custom tables.*

Your next step is to get a sense of your table dimensions. You'll ask:

-   Will each fit neatly into a portrait orientation, or should the page be rotated 90 degrees into landscape view?
-   Even in landscape view, is the table too wide to be shown properly, and must be split into smaller tables by column?
-   Is the table too long for one page, and must be split into smaller tables by row?

This section contains the steps to answer these questions.

> **NOTE**: This workflow is reflected in the create_tables_doc.R file and its associated functions (such as `render_lg_table()`, `ID_tbl_length_class()`, `ID_tbl_width_class()`, and `export_split_tbls()`).

Run this code:

```{r eval=F}
my_plot_name <- "example_tab.rda" # replace this with the filename of your table rda
tables_dir <- getwd() # replace this with the location of your tables folder

# Identify table width class
tbl_orient <- ID_tbl_width_class(
  plot_name = my_plot_name,
  tables_dir = tables_dir,
  portrait_pg_width = 5 # 5 inches = the threshold for maximum table width before it needs to be resized, rotated, and/or split
)

# Identify table length class
tbl_length <- ID_tbl_length_class(
      plot_name = my_plot_name,
      tables_dir = tables_dir
    )

table_specs <- list(tbl_orient, tbl_length)

# Identify table dimension class (length x width)
tbl_class <- dplyr::case_when(
      table_specs[[1]] == "regular" & table_specs[[2]] == "regular" ~ "reg_reg",
      table_specs[[1]] == "regular" & table_specs[[2]] == "long" ~ "reg_long",
      table_specs[[1]] == "wide" & table_specs[[2]] == "regular" ~ "wide_reg",
      table_specs[[1]] == "wide" & table_specs[[2]] == "long" ~ "wide_long",
      table_specs[[1]] == "extra-wide" & table_specs[[2]] == "regular" ~ "ewide_reg",
      table_specs[[1]] == "extra-wide" & table_specs[[2]] == "long" ~ "ewide_long",
      TRUE ~ "unknown"
    )

max_rows <- ifelse(tbl_orient == "regular", 38, 28)

print(paste0("Your table dimension is ", tbl_class, ". The maximum number of table rows per page is ", max_rows, "."))

tables_path <- fs::path(
    tables_dir,
    "tables",
    paste0(plot_name, "_table.rda")
  )

load(tables_path)

# Get table length in rows
split_table_rows <- length(rda[[1]]$`_data`[[1]])
split_tables_rowwise <- ceiling(split_table_rows / max_rows)

if (tbl_length == "long") {
  print(paste0("Your table must be split row-wise. Each table can have, at most, ", max_rows, "rows."))
}
if (tbl_orient == "extra-wide") {
  print(paste0("Your table must be split column-wise. Proceed to the section 'Table splitting' (regarding extra-wide tables) to identify the number of smaller tables your original table must be split into."))
}
```

You should see one of six outputs: reg_reg, reg_long, wide_reg, wide_long, ewide_reg, or ewide_long.

|  |  |  |  |  |  |  |
|-----------|:---------:|:---------:|:---------:|-----------|-----------|-----------|
| **Classification** | reg_reg | reg_long | wide_reg | wide_long | ewide_reg | ewide_long |
| **Dimensions** | Regular width & length | Regular width, long | Wide, regular length | Wide, long | Extra-wide, regular length | Extra-wide, long |
| **Table width** | \< 5" wide | \< 5" wide | at least 5" and up to 12" wide | at least 5" and up to 12" wide | \>12" wide | \>12" wide |
| **Table length**[^id_width-1] | \<= 38 rows | \> 38 rows | \<= 28 rows | \> 28 rows | \<= 28 rows | \> 28 rows |

: Table width and length classification system coded in the `create_tables_doc` function.

[^id_width-1]: The maximum number of rows drops from 38 (for regular-width tables) to 28 (for wide or extra-wide tables) because the latter two tables are rotated on a landscape page, which offers less space for table rows.

**Regular-length tables**

*If your output is reg_reg*, you can display your table in a .qmd chunk without any additional steps. Jump to section [Add custom tables and figures](#sec-add-plots) and proceed.

*If your output is wide_reg*, you will place your table's .qmd chunk in special braces that change the page orientation to landscape. Jump to section [Add custom tables and figures](#sec-add-plots) and proceed.

*If your output is ewide_reg*, you will place your table's .qmd chunk in special braces that change the page orientation to landscape *and* split it into narrower tables, each of which are displayed on separate pages. Proceed to the section [Table splitting (for extra-wide tables only)](#sec-table-splitting).

**Long tables**

*If your output is reg_long*, you will split your table into shorter tables, each of which are displayed on separate pages. Proceed to the section \_\_\_\_.

*If your output is wide_long*, you will place your table's .qmd chunk in special braces that change the page orientation to landscape *and* split your table into shorter tables, each of which are displayed on separate pages. Proceed to the section \_\_\_\_\_\_.

*If your output is ewide_long*, you will place your table's .qmd chunk in special braces that change the page orientation to landscape; split it into narrower tables, each of which are displayed on separate pages; *and* split your table into shorter tables, each of which are displayed on separate pages. Proceed to the section \_\_\_\_\_.

#### Table splitting

##### Extra-wide tables

Extra-wide tables will need to be split among multiple pages. Run this code to learn the number of tables into which your original table will be split (by column) *and* export the split tables (as a list) into a new rda:

```{r eval=F}
my_plot_name <- "example_tab.rda" # replace this with the filename of your table rda
tables_dir <- getwd() # replace this with the location of your tables folder
essential_cols <- 1:2 # replace this with the columns that will be retained between the split tables, formatted as a sequence (e.g., 1:2 for columns 1-2, or 1 for a single column)

export_split_tbls(
  tables_dir = tables_dir,
  plot_name = my_plot_name,
  essential_columns = essential_cols
)
```

If the number returned was, say, 5, then since each split table must be in its own chunk, you will need 5 chunks to show your split tables by column (if your table is *ewide-long*, then you must also split tables by row). You will also need an additional chunk to load in the data. Each table needs its own chunk, and therefore label, so that it can be cross-referenced in text.

You can check out your split tables in the new rda saved in your tables_dir folder. It will have the same filename as your original rda except that it will contain "\_split" in the filename. For instance, split tables created from "example_tab.rda" will be saved in "example_tab_table_split.rda".

##### Long tables

Long tables will need to be split among multiple pages. Refer to the code chunk above to learn the number of tables into which your original table will be split by row.

If the number returned was, say, 3, then since each split table must be in its own chunk, you will need 3 chunks to show your split tables by row. You will also need an additional chunk to load in the data. Each table needs its own chunk, and therefore label, so that it can be cross-referenced in text.

##### Extra-wide, long tables

Extra-wide, long tables will need to be split by column *and* row. Read both "Extra-wide tables" and "Long tables" sections above, then proceed to the next section for guidance.
