### Add custom tables and figures {#sec-add-plots_coding}

In your 'report' folder, open your figures and tables docs (probably named something like '08_tables.qmd' and '09_figures.qmd'). Add the following code there.

#### Tables

Follow the instructions in the appropriate section below based on your table classification: reg_reg, reg_long, wide_reg, wide_long, ewide_reg, or ewide_long.

Regardless of the table type, you must **first** include a setup chunk to load your data and split the caption into usable objects:

````{verbatim}
```{r}
#| label: 'tab-setup'
#| include: false

# Load the rda
load(file.path(tables_dir, 'example_tab_table.rda'))

# Save to specific objects for the display chunks
your_table <- rda$table
your_cap <- rda$caption
```
````

-----

##### Regular tables (reg_reg)

*Portrait orientation, fits on one page.*

Add this chunk to display the table:

````{verbatim}
```{r}
#| label: 'tbl-custom'
#| echo: false
#| warning: false
#| tbl-cap: !expr your_cap

your_table |>
  gt::cols_width(everything() ~ pct(20))
```
````

##### Wide tables (wide_reg)

*Landscape orientation, fits on one page.*

Wrap the chunk in landscape braces:

````{verbatim}
::: {.landscape}

```{r}
#| label: 'tbl-custom-wide'
#| echo: false
#| warning: false
#| tbl-cap: !expr your_cap

your_table |>
  gt::tab_options(table.width = pct(100), table.layout = 'auto') |>
  gt::cols_width(everything() ~ pct(20))
```

:::
````

##### Long tables (reg_long or wide_long)

*Splits by row across multiple pages. wide_long uses landscape; reg_long uses portrait.*

You will need a separate chunk for each page of the split (e.g., if `split_tables_rowwise` was 2, repeat this twice, changing the index in `grp_pull`).

> **Note**: For wide_long, wrap the chunk in `::: {.landscape}` braces, as shown for wide_reg tables, above.

````{verbatim}
```{r}
#| label: 'tbl-custom-long1'
#| echo: false
#| tbl-cap: !expr paste0(your_cap, ' (1 of 2)')

your_table |>
  gt::tab_options(table.width = pct(100), table.layout = 'auto') |>
  gt::cols_width(everything() ~ pct(20)) |> 
  asar::gt_split(row_every_n = 38) |> # Use 28 for wide_long
  gt::grp_pull(1)                    # Change this index for subsequent chunks
```
````

##### Extra-wide tables (ewide_reg)

*Landscape orientation, split by column across multiple pages.*

First, load the split rda:

````{verbatim}
```{r}
#| label: 'tbl-setup-split'
#| include: false
load(file.path(tables_dir, 'example_tab_table_split.rda'))
table_list_split <- table_list
```
````

Then, for each split table in the list, create a chunk:

````{verbatim}
::: {.landscape}

```{r}
#| label: 'tbl-custom-ewide1'
#| echo: false
#| tbl-cap: !expr paste0(your_cap, ' (1 of 3)')

table_list_split[[1]] |>
  gt::tab_options(table.width = pct(100), table.layout = 'auto') |>
  gt::cols_width(everything() ~ pct(20))
```

:::
````

##### Extra-wide, long tables (ewide_long)

*Landscape orientation, split by both column AND row.*

This requires a nested approach. You must create chunks for every combination of column-split and row-split.

````{verbatim}
::: {.landscape}

```{r}
#| label: 'tbl-custom-ewide-long1-1'
#| echo: false
#| tbl-cap: !expr paste0(your_cap, ' (Column split 1 of 2, Row split 1 of 3)')

table_list_split[[1]] |>
  gt::tab_options(table.width = pct(100), table.layout = 'auto') |>
  gt::cols_width(everything() ~ pct(20)) |>
  asar::gt_split(row_every_n = 28) |>
  gt::grp_pull(1)
```

:::
````

##### Final step: page break

If you are adding multiple tables, insert a page break between them:

```{verbatim}
{{< pagebreak >}}
```

Proceed to _________

#### Figures

The workflow for figures is very similar to the workflow for regular tables.

In your figures doc, add the following chunk to display the figure in your .qmd when it's rendered:

````{verbatim}
```{r} 
#| label: 'fig-custom1' # choose a label
#| echo: false
#| warning: false
#| fig-cap: !expr This is your figure caption.
#| fig-alt: !expr This is your figure alternative text.
# Import your data
your_df <- data.frame(
  "x" = c(1, 2, 3),
  "y" = c(4, 5, 6)
)

# Make your figure
your_figure <- ggplot2::ggplot(
  your_df,
  aes(
    x = x,
    y = y
  )
) +
  ggplot2::geom_point()

# Show your figure
your_figure
``` 
````

If you are adding another figure, add a page break first:

```{verbatim}
{{< pagebreak >}}
```

You're almost done! Proceed to the last section ([Update report skeleton](#sec-update-skeleton)) to finish adding your custom figures to your report.
