### Set up Quarto markdown (.qmd) files to display custom tables and figures {#sec-set-up-qmd}

You will create one .qmd file for your tables and one .qmd file for your figures.

#### Set up templates

Open a blank .qmd script (in RStudio, click File -\> New File -\> Quarto Document -\> Create Empty Document (button in the bottom left corner of the popup)).

Remove all text from the new document.

Switch from Visual to Source mode ([instructions here](https://quarto.org/docs/visual-editor/#switching-modes)).

Save this new file with a clear name in a logical location. We recommend filenames like 08_tables_custom.qmd and 09_figures_custom.qmd, then saving these files in the same report folder as the rest of your report files. We include separate workflows for tables and figures, below.

#### Tables

In your tables doc, add your first R chunk (shown below). This chunk is used to establish your tables_dir folder and load the library needed to plot your tables (`gt`).

Click the following links to learn more about [cross-references](https://quarto.org/docs/computations/execution-options.html) (label) and [chunk options](https://quarto.org/docs/computations/execution-options.html) (echo, warning, etc.).

````{verbatim}
```{r} 
#| label: 'set-rda-dir-tbls-custom' # you can set this label as you wish, as long as it's not repeated in any other R chunk in, for instance, the 08_tables.qmd or 09_figures.qmd files. Labels are used for cross-referencing in your report text.
#| echo: false
#| warning: false
#| include: false
library(gt)
tables_dir <- "C:/path/to/your/tables" # set the path to your tables_dir
``` 
````

Now, follow the instructions in the appropriate section depending on your table width: regular, wide, or extra wide.

##### Regular table

Add the following chunk to your script to import your table's rda and save objects containing the table and caption:

````{verbatim}
```{r} 
#| label: 'tbl-setup-custom1' # choose a label
#| echo: false
#| warning: false
#| include: false
# load your rda and save with a table-specific name, then delete "rda"
load(file.path(tables_dir, "example_tab.rda"))
example_table_rda <- rda
rm(rda)

# save table, caption as separate objects
example_table <- example_table_rda$table
example_cap <- example_table_rda$caption
``` 
````

Then, add this chunk to display the table in your .qmd when it's rendered:

````{verbatim}
```{r} 
#| label: 'tbl-custom1' # choose a label
#| echo: false
#| warning: false
#| tbl-cap: !expr example_cap
example_table
``` 
````

If you are adding another table, regardless of size, add a page break first:

```{verbatim}
{{< pagebreak >}}
```

Keep adding as many tables as you'd like. Keep in mind that each table must have a chunk that imports it, and another chunk that displays it, as shown above.

You're almost done! Proceed to the last section ([Update report skeleton](#sec-update-skeleton)) to finish adding your custom tables to your report.

##### Wide table

Add the following chunk to your script to import your table's rda and save objects containing the table and caption:

````{verbatim}
```{r} 
#| label: 'tbl-setup-custom1' # choose a label
#| echo: false
#| warning: false
#| include: false
# load your rda and save with a table-specific name, then delete "rda"
load(file.path(tables_dir, "example_tab.rda"))
example_table_rda <- rda
rm(rda)

# save table, caption as separate objects
example_table <- example_table_rda$table
example_cap <- example_table_rda$caption
``` 
````

Before you add your next chunk, add [landscape braces](https://quarto.org/docs/authoring/article-layout.html#landscape-mode) to ensure your table will be displayed on a landscape-oriented page:

```{verbatim}
::: {.landscape}
```

Then, add this chunk to display the table in your .qmd when it's rendered:

````{verbatim}
```{r} 
#| label: 'tbl-custom1' # choose a label
#| echo: false
#| warning: false
#| tbl-cap: !expr example_cap

# the tab_options() and cols_width() lines format the table to fit the landscape page
example_table |>
  gt::tab_options(
    table.width = pct(100),
    table.layout = 'auto'
  ) |>
  gt::cols_width(
    everything() ~ pct(20)
  )
``` 
````

And add three colons to close the landscape braces:

```{verbatim}
:::
```

> **NOTE**: Landscape braces only work in [Quarto v1.6+](https://quarto.org/docs/blog/posts/2024-11-25-1.6-release/). Please read [this Discussion](https://github.com/nmfs-ost/asar/discussions/186) to learn how to check your Quarto version and update it if necessary.

If you are adding another table, regardless of size, add a page break first:

```{verbatim}
{{< pagebreak >}}
```

Keep adding as many tables as you'd like. Keep in mind that each table must have a chunk that imports it, and another chunk that displays it, as shown above.

You're almost done! Proceed to the last section ([Update report skeleton](#sec-update-skeleton)) to finish adding your custom tables to your report.

##### Extra-wide table

Add the following chunk to your script to:

-   import your *original table's* rda
-   save original table's caption as a separate object
-   import your *split tables'* rda
-   save split tables' caption specifiers as a separate object

````{verbatim}
```{r} 
#| label: 'tbl-labels-custom1' # choose a label
#| echo: false
#| warning: false
#| include: false
# load your original table rda and save with a table-specific name, then delete "rda"
load(file.path(tables_dir, "example_tab.rda"))
example_table_rda <- rda
rm(rda)

# save caption as separate object
example_cap <- example_table_rda$caption

# load split table rda and save with a table-specific name, then delete "table_list"
load(file.path(tables_dir, "example_table_split.rda"))
example_table_rda_split <- table_list
rm(table_list)

# extract table caption specifiers
example_cap_split <- names(example_table_rda_split)
``` 
````

Example caption specifiers could include fleets only shown in specific tables.

Before you add your next chunk, add [landscape braces](https://quarto.org/docs/authoring/article-layout.html#landscape-mode) to ensure your table will be displayed on a landscape-oriented page:

```{verbatim}
::: {.landscape}
```

Now, you'll add as many chunks as you have split tables. Here's what the first one will look like:

````{verbatim}
```{r} 
#| label: 'tbl-custom1' # choose a label
#| echo: false
#| warning: false
#| tbl-cap: !expr paste0(example_cap, '(', example_cap_split[[1]], ')') # this pastes the original table caption with the specifiers associated with table 1

# the tab_options() and cols_width() lines format the table to fit the landscape page
# plot split table 1
example_table_rda_split[[1]] |>
  gt::tab_options(
    table.width = pct(100),
    table.layout = 'auto'
  ) |>
  gt::cols_width(
    everything() ~ pct(20)
  )
``` 
````

Copy and paste this chunk for as many split tables as you have. Then, replace the instances of "1" with the split table number (e.g., 2,3,4, etc.). In the example above, "1" occurs in the chunk options (label, tbl-cap) and the table-plotting code itself.

Now, add three colons to close the landscape braces:

```{verbatim}
:::
```

> **NOTE**: Landscape braces only work in [Quarto v1.6+](https://quarto.org/docs/blog/posts/2024-11-25-1.6-release/). Please read [this Discussion](https://github.com/nmfs-ost/asar/discussions/186) to learn how to check your Quarto version and update it if necessary.

Keep adding as many tables as you'd like. Keep in mind that each table must have a chunk that imports it, and another chunk that displays it, as shown above.

You're almost done! Proceed to the last section ([Update report skeleton](#sec-update-skeleton)) to finish adding your custom tables to your report.

#### Figures

The workflow for figures is very similar to the workflow for regular tables.

In your figures doc, add your first R chunk (shown below). This chunk is used to establish your figures_dir folder.

Click the following links to learn more about [cross-references](https://quarto.org/docs/computations/execution-options.html) (label) and [chunk options](https://quarto.org/docs/computations/execution-options.html) (echo, warning, etc.).

````{verbatim}
```{r} 
#| label: 'set-rda-dir-figs-custom' # you can set this label as you wish, as long as it's not repeated in any other R chunk in, for instance, the 08_tables.qmd or 09_figures.qmd files. Labels are used for cross-referencing in your report text.
#| echo: false
#| warning: false
figures_dir <- "C:/path/to/your/figures" # set the path to your figures_dir
``` 
````

Add the following chunk to your script to import your figure's rda and save objects containing the figure, caption, and alternative text:

````{verbatim}
```{r} 
#| label: 'fig-setup-custom1' # choose a label
#| echo: false
#| warning: false
#| include: false
# load your rda and save with a figure-specific name, then delete "rda"
load(file.path(figures_dir, "example_fig.rda"))
example_figure_rda <- rda
rm(rda)

# save figure, caption, and alternative text as separate objects
example_plot <- example_figure_rda$figure
example_cap <- example_figure_rda$caption
example_alt_text <- example_figure_rda$alt_text
``` 
````

Then, add this chunk to display the figure in your .qmd when it's rendered:

````{verbatim}
```{r} 
#| label: 'fig-custom1' # choose a label
#| echo: false
#| warning: false
#| fig-cap: !expr example_cap
#| fig-alt: !expr example_alt_text
example_plot
``` 
````

If you are adding another figure, add a page break first:

```{verbatim}
{{< pagebreak >}}
```

You're almost done! Proceed to the last section ([Update report skeleton](#sec-update-skeleton)) to finish adding your custom figures to your report.
