### Set up Quarto markdown (.qmd) files to display custom tables and figures {#sec-set-up-qmd}

You will create one .qmd file for your tables and one .qmd file for your figures.

#### Set up templates

Open a blank .qmd script (in RStudio, click File -\> New File -\> Quarto Document -\> Create Empty Document (button in the bottom left corner of the popup)).

Remove all text from the new document.

Switch from Visual to Source mode ([instructions here](https://quarto.org/docs/visual-editor/#switching-modes)).

Save this new file with a clear name in a logical location. We recommend filenames like 08_tables_custom.qmd and 09_figures_custom.qmd, then saving these files in the same report folder as the rest of your report files. We include separate workflows for tables and figures, below.

#### Tables

Follow the instructions in the appropriate section depending on your table width: regular or wide.

##### Regular table

Add the following chunk to display the table in your .qmd when it's rendered:

````{verbatim}
```{r} 
#| label: 'tbl-custom1' # choose a label
#| echo: false
#| warning: false
#| tbl-cap: !expr This is your table caption.

# Import your data
your_df <- data.frame(
  "x" = c(1, 2, 3),
  "y" = c(4, 5, 6)
)

# Make your table
your_table <- gt::gt(your_df)

# Show your table
your_table
``` 
````

If you are adding another table, regardless of size, add a page break first:

```{verbatim}
{{< pagebreak >}}
```

Keep adding as many tables as you'd like. Keep in mind that each table must have a chunk that imports it, and another chunk that displays it, as shown above.

You're almost done! Proceed to the last section ([Update report skeleton](#sec-update-skeleton)) to finish adding your custom tables to your report.

##### Wide table

Before you add your first chunk, add [landscape braces](https://quarto.org/docs/authoring/article-layout.html#landscape-mode) to ensure your table will be displayed on a landscape-oriented page:

```{verbatim}
::: {.landscape}
```

Then, add the following chunk to display the table in your .qmd when it's rendered:

````{verbatim}
```{r} 
#| label: 'tbl-custom1' # choose a label
#| echo: false
#| warning: false
#| tbl-cap: !expr This is your table caption.

# Import your data
your_df <- data.frame(
  "x" = c(1, 2, 3),
  "y" = c(4, 5, 6)
)

# Make your table
your_table <- gt::gt(your_df) |>
  flextable::fit_to_width(max_width = 8) # this extra line ensures the table will be resized to a maximum of 8" wide

# Show your table
your_table
``` 
````

And add three colons to close the landscape braces:

```{verbatim}
:::
```

> **NOTE**: Landscape braces only work in [Quarto v1.6+](https://quarto.org/docs/blog/posts/2024-11-25-1.6-release/). Please read [this Discussion](https://github.com/nmfs-ost/asar/discussions/186) to learn how to check your Quarto version and update it if necessary.

If you are adding another table, regardless of size, add a page break first:

```{verbatim}
{{< pagebreak >}}
```

Keep adding as many tables as you'd like. Keep in mind that each table must have a chunk that imports it, and another chunk that displays it, as shown above.

You're almost done! Proceed to the last section ([Update report skeleton](#sec-update-skeleton)) to finish adding your custom tables to your report.

#### Figures

The workflow for figures is very similar to the workflow for regular tables.

In your figures doc, add the following chunk to display the figure in your .qmd when it's rendered:

````{verbatim}
```{r} 
#| label: 'fig-custom1' # choose a label
#| echo: false
#| warning: false
#| fig-cap: !expr This is your figure caption.
#| fig-alt: !expr This is your figure alternative text.
# Import your data
your_df <- data.frame(
  "x" = c(1, 2, 3),
  "y" = c(4, 5, 6)
)

# Make your figure
your_figure <- ggplot2::ggplot(
  your_df,
  aes(
    x = x,
    y = y
  )
) +
  ggplot2::geom_point()

# Show your figure
your_figure
``` 
````

If you are adding another figure, add a page break first:

```{verbatim}
{{< pagebreak >}}
```

You're almost done! Proceed to the last section ([Update report skeleton](#sec-update-skeleton)) to finish adding your custom figures to your report.
