### Update report skeleton {#sec-update-skeleton}

Finally, update the report skeleton so that it imports your new custom table and figure .qmd files.

Open your skeleton .qmd file. Scroll down to the chunks that contain tables and figures (their labels are "tables" and "figures", respectively).

Your "tables" chunk will look similar, if not identical, to this:

````{verbatim}
```{r, results='asis'}
#| label: 'tables'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child("08_tables.qmd", quiet = TRUE)
cat(a, sep = "\n")
```
````

If you are adding custom tables, copy the two lines of code that start with "a \<- knitr::" and "cat". Paste them below the original two lines of code in the same chunk. Then, in those two new lines, make the following changes:

1.  Change "a" to "b"
2.  Replace '08_tables.qmd' with the name of your new custom .qmd

For example:

````{verbatim}
```{r, results='asis'}
#| label: 'tables'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child("08_tables.qmd", quiet = TRUE)
cat(a, sep = "\n")
b <- knitr::knit_child("08_tables_custom.qmd", quiet = TRUE)
cat(b, sep = "\n")
```
````

If you're adding custom figures, follow the same steps with the figures chunk.

Save the changes to your skeleton, and you're done! Your custom tables and figures will be included in your report.
